A comparative analysis of penalized linear mixed models in structured genetic data

### 2020-10-15 main.tex
Small typos

Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$. We see this by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of the projection matrix of $\bX$.  Let $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$ denote the projection matrix of $\bX$, with $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denoting the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively. Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\bX \boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.  This in turn means that
\begin{align}
    \label{eqn:partition_e}    \by &= \bX\bbeta + \bZ\bgamma + \beps \notag \\
    &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps \notag\\
    &=  \bX \bbeta{^*} + \beps{^*};
\end{align}
note that $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat^*$ and $\boldsymbol{\psi}$ is absorbed by the residual error term. In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.  This result provides the explicit mathematical justification for our statement from Section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.

@@

In Section \ref{sec:lmmlasso} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. Both methods perform quite similarly in terms of $\bbeta$ MSE and variable selection. Becasue of this, the preceding results, though generated using the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods in the $n = 200, p = 1000$ setting. The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding. However, given the generally strong performance of LMM-Lasso, we also wish to discuss the key differences between these two implementations and provide guidance for potential users.

>>>>

Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$. We see this by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of the projection matrix of $\bX$.  Let $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$ denote the projection matrix of $\bX$, with $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denoting the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively. Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.  This in turn means that
\begin{align}
    \label{eqn:partition_e}    \by &= \bX\bbeta + \bZ\bgamma + \beps \notag \\
    &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps \notag\\
    &=  \bX \bbeta{^*} + \beps{^*};
\end{align}
note that $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat^*$ and $\boldsymbol{\psi}$ is absorbed by the residual error term. In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.  This result provides the explicit mathematical justification for our statement from Section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.

@@

In Section \ref{sec:lmmlasso} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. Both methods perform quite similarly in terms of $\bbeta$ MSE and variable selection. Because of this, the preceding results, though generated using the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods in the $n = 200, p = 1000$ setting. The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding. However, given the generally strong performance of LMM-Lasso, we also wish to discuss the key differences between these two implementations and provide guidance for potential users.

### 2020-10-15 supplement.tex
Small typos

% 1D LINEAR ADMIXTURE KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/admixed_kinship.png}
    \caption{1D Linear Admixture with coarse subpopulation structure (50 subjects per subpopulation), estimated kinship matrix.}
    \label{fig:admixed}
\end{figure}

% INDEPENDENT SUBPOPULATIONS, COARSE, KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/indep_coarse_kinship.png}
    \caption{Independent Subpopulations with coarse subpopulation structure (50 subjects per subpopulation), estimated kinship matrix.}
    \label{fig:indep_coarse}
\end{figure}

% INDEPENDENT SUBPOPULATIONS, FINE, KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/indep_fine_kinship.png}
    \caption{Independent Subpopulations with subpopulation structure (2 subjects per subpopulation), estimated kinship matrix.}
    \label{fig:indep_fine}
\end{figure}

% EMPIRICAL KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/empirical_kinship.png}
    \caption{Empirical estimated kinship matrix.}
    \label{fig:empirical}
\end{figure}

>>>>

% 1D LINEAR ADMIXTURE KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/admixed_kinship.png}
    \caption{1D Linear Admixture with coarse subpopulation structure (50 subjects per subpopulation), RRM.}
    \label{fig:admixed}
\end{figure}

% INDEPENDENT SUBPOPULATIONS, COARSE, KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/indep_coarse_kinship.png}
    \caption{Independent Subpopulations with coarse subpopulation structure (50 subjects per subpopulation), RRM.}
    \label{fig:indep_coarse}
\end{figure}

% INDEPENDENT SUBPOPULATIONS, FINE, KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/indep_fine_kinship.png}
    \caption{Independent Subpopulations with fine subpopulation structure (2 subjects per subpopulation), RRM.}
    \label{fig:indep_fine}
\end{figure}

% EMPIRICAL KINSHIP
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/empirical_kinship.png}
    \caption{Empirical RRM.}
    \label{fig:empirical}
\end{figure}
